---
title: "Gettin' groovy with Bookdown"
author: "CB Wall -- cbwall@hawaii.edu"
date: "11/25/2020"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2: default
linkcolor: blue
header-includes: 
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---
```{r setup chunk, setup, include = FALSE, cache=FALSE, message=FALSE, warning=FALSE}
if (!require('knitr')) install.packages('knitr'); library('knitr')
knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.align='center')

# load packages
if (!require("pacman")) install.packages("pacman") # for rapid install if not in library

# use pacman to load all the packages you are missing!
pacman::p_load('knitr', 'lme4', 'lmerTest', 'effects', 'dplyr', 'plotrix', 'car', 'bookdown')
```
# Main text {-}
 
**Git up! BookDowwwwwn!**  
For the advanced Markdown-ers, you can also write and code a manuscript in markdown that can be exported to Word, with references embedded. Markdown magic! [R markdown: Publsihing workflows and manuscript.](https://danovando.github.io/publications-with-rmarkdown/presentations/pubs-with-rmarkdown)
  
I've found this does not work for html--which is fine--but the pdf and word outputs seem to flow quite well. Make sure to load the package `bookdown` and set the options for output at the top of the Markdown script as I have above-- with   
`output:`  
  `bookdown::word_document2: default`  
  `bookdown::pdf_document2: default`   
  `bookdown::html_document2: default`  
  
Let's try to embed a figure in the document we are making now. If I wanted the reference to the figure to go chronological in the code (1 to the nth) and I name the code chunk accordingly, we can 'call the chunk' using a reference!   
  
This uses the package `bookdown` and this approach is called `cross referencing`. It works by typing `\@ref(ref-type:chunk-name)`, which produces a linked number to the referenced object. Let's try one, here using *ref-type* = "fig" and the *code chunk name* = "test-fig".  **Note** the code chunk name must not have spaces. Let's start a new page and call the code chunk

\newpage
  
**Chlorophyll?! More like bore-o-phyll! (Fig.\@ref(fig:test-plot)).**
```{r test-plot, fig.cap = "Chlorophyll *a* boxplot", fig.dim=c(5,3), fig.align='center'}
data<-read.csv("data/coral_data.csv")
hist(data$chla, col="seagreen", freq=F, main="Chlorophylls in coral symbionts", ylab="density",
      xlab=expression(paste("Chlorophyll", ~italic("a"+"c"[2]), ~(mu*g~cm^-2), sep="")))
par(new=T)
lines(density(data$chla),lwd=3,col="mediumseagreen")
```
  
Notice the text updates to show *Figure 1* and the code chunk (displayed above) produced the figure and adds the *Figure 1* to the caption, despite the fact that I did not directly name it Figure 1 using `fig.cap`. This is great because now R will reference these figures chronologically and link the # to the code chunk name. If a reviewer says, "I think this should be moved to the supplement and have a new Figure 1," we can do that easily enough.
  
Let's try... making another figure, this one will be figure 2--as per reviewer requests

*We have appeased the reviewer with this fatted calf and a blood offering. We hope this new data with satiate the appetite of the reviewer*.  Please see (Fig.\@ref(fig:biomass)), which is awesome, and chlorophyll data now in the supplement.
```{r biomass, results="hide", fig.dim=c(5,3), fig.align='center', fig.cap="Coral tissue biomass", echo=FALSE}
# use results="hide" here to suppress messages associate with "dev.off" and dev.print to save figure

MC.df<-data[(data$Species=="MC"),] ## Montipora capitata alone
PC.df<-data[(data$Species=="PC"),] ## Porites compress alone

par(mar = c(4, 4, 3, 0.1))

plot(density(MC.df$biomass), lwd=2, col="darksalmon", xlim=c(0,100), main="Tissue biomass in two coral species", cex.main=0.8, xlab=expression(paste("Biomass", ~(mg~cm^-2), sep="")))
lines(density(PC.df$biomass), lwd=2, col="darkslategray3", yaxt="n", ylab="", xaxt="n", xlab="")
legend("topright", legend=c(expression(italic("Montipora capitata")), expression(italic("Porites compressa"))), lwd=2, col=c("darksalmon", "darkslategray3"), bty="n")


```


\newpage
# Supplementary material {-}

\beginsupplement
Here is the chlorophyll data in a new section, with numbers restarted. We accomplished this by using the `header` information in the top of the script.  (Fig.\@ref(fig:test-plot2)).
```{r test-plot2, fig.cap = "Chlorophyll *a* boxplot", fig.dim=c(5,3), fig.align='center', echo=FALSE}
data<-read.csv("data/coral_data.csv")
hist(data$chla, col="seagreen", freq=F, main="Chlorophylls in coral symbionts", ylab="density",
      xlab=expression(paste("Chlorophyll", ~italic("a"+"c"[2]), ~(mu*g~cm^-2), sep="")))
par(new=T)
lines(density(data$chla),lwd=3,col="mediumseagreen")
```
  
  
If you knitted this for the first time, looked at the output, and  said "Holy SH*T!"... I did too!  
Now a warning, if you knit the above with the `header-includes` segment, this may only export the way you wish for pdfs. So if exporting to word is what you want to do, there are some caveates--which might mean moving the order of chunks around so that they can be listed chronologically. BUT that is doable.  You can see more on this at [Stack Overflow](https://stackoverflow.com/questions/50223141/rmarkdown-bookdown-separate-figure-numbering-for-supplemental-section), where I got the `header-includes` info...